---
title: 'Put your title here'
author: Brian C. Stock^1^, Coauthor Num1^2^, Coauther Num2^3^, Your Adviser^1^
date: ''
output:
  pdf_document:
    fig_caption: yes
    includes:
      in_header: options.sty
    latex_engine: xelatex
csl: ices-journal-of-marine-science.csl
bibliography: mypapers.bib
---

```{r setup, include=FALSE, echo=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=FALSE, tidy.opts=list(width.cutoff=60), warning = FALSE, message = FALSE)
# knitr::opts_knit$set(root.dir = '../' ) # set directory one level up from this file
library(knitr)
library(tidyverse)
library(ggsidekick)
library(reshape2)
library(viridis)
library(date)
library(gridExtra)
library(pander)
library(kableExtra)
library(png)
library(forcats)
```

^1^phone number, email, Scripps Institution of Oceanography, University of California, San Diego, La Jolla, California
92093 USA\
^2^Coauthor address\
^3^Coauthor address\


## Abstract

Abstract text

## Keywords

Keyword 1, keyword 2, ...

## Introduction  
  
Write your paper like normal. Add citations like this [@nmfs2016bycatch]. Multiple citations like so [@thorson2017; @ovaskainen2016]. A more complex reference [for details see NWFSC -@nmfs2016bycatch].

Sometimes you want to put in equations inline, 
$\sum_{ s=1 }^{ S }{ \frac { { d }_{ s } }{ { r }_{ s } }  } { R }_{ s }$, where ${ d }_{ s }$ is something you care about, and *s*, ${ r }_{ s }$ is something else.

Super important equations get their own lines like so:
$$ log(\text{Bycatch}) = \alpha + \beta \ log(\text{Effort}) + \epsilon$$
$$ \epsilon \sim \mathcal{N}(0,\,\sigma^{2})$$

## Methods

### Sub-heading

We used data from this region (Fig. \ref{fig:effort}). Here's a list of species we care about (Table \ref{tab:species-list}), and see Table \ref{tab:species-list-byyear} for a list by year.

### Sub-heading 2

More text and equations.

## Results

### Our big main result

We found some results (Figs. \ref{fig:effort} and \ref{fig:model-comparison}).

### Our slightly more subtle result

The main result was cool, but then there is some nuance to show (27% lower averaged across species, Fig. \ref{fig:model-comparison}).

## Discussion

Wow those results were interesting!

## Acknowledgements

Funding from XXXX. Thanks to YYYY.  

## References

<div id="refs"></div>

\pagebreak

```{r species-list, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
allyears = readRDS("figures/allyears.rds")
      
kable(allyears, "latex", booktabs = T, caption="\\label{tab:species-list}Total bycatch (mt) and bycatch rate (percent of hauls) for species selected from the U.S. West Coast Groundfish Observer Program (WCGOP) dataset. All selected species are exclusively discarded. The summarized data are 35,440 post-IFQ hauls (4,007 trips) observed from 2011-2015 in the area north of Cape Falcon, Oregon (45.77° N).") %>%
  kable_styling(latex_options=c("basic"))
```

\pagebreak

```{r species-list-byyear, echo=FALSE, message=FALSE, warnings=FALSE, results='asis'}
byyear = readRDS("figures/byyear.rds")
      
kable(byyear, "latex", booktabs = T, caption="\\label{tab:species-list-byyear}Total bycatch (mt) and bycatch rate (percent of hauls) for species selected from the U.S. West Coast Groundfish Observer Program (WCGOP) dataset. All selected species are exclusively discarded. The summarized data are 35,440 post-IFQ hauls (4,007 trips) observed from 2011-2015 in the area north of Cape Falcon, Oregon (45.77° N).") %>%
  kable_styling(latex_options=c("basic")) %>% 
  landscape() %>%
  add_header_above(c(" "=1, "2011"=2, "2012"=2, "2013"=2, "2014"=2, "2015"=2))
```

\pagebreak

<!-- Figure 1, uses raw data so code not included here -->
<!-- https://www.zevross.com/blog/2017/06/19/tips-and-tricks-for-working-with-images-and-figures-in-r-markdown-documents/ -->
```{r effort, echo=FALSE, message=FALSE, warnings=FALSE, out.width="2.5in", fig.align = "center", fig.cap="Fishing effort density in the West Coast groundfish trawl fishery from 2011 to 2015 in the area north of Cape Falcon, Oregon (45.77° N). The West Coast Groundfish Observer Program monitored and collected data from 35,440 hauls from all (100 percent) of the 4,007 trips. Fishing effort was smoothed using a bivariate kernel density estimate ('bkde2D' function in R package 'KernSmooth') to ensure that fishing locations were anonymized."}
# fig1_path <- "../figures/fig1_premade.png"
fig1_path <- "figures/fig1_premade.png"
knitr::include_graphics(fig1_path)
```

\pagebreak

```{r model-comparison, echo=FALSE, message=FALSE, warnings=FALSE, fig.height=6, fig.width=7, fig.align = "center", fig.cap="Put your caption here. Watch out for symbols like percent signs... I couldn't figure out how to add them!"}
dat = readRDS("figures/results_summary.rds")
dat <- filter(dat, !(model %in% c("RF Cubist","RF Xu2","RF Xu4")))
dat$model <- as.character(dat$model)

# for now we'll only plot the simulations with 20% coverage
dat = filter(dat, pct_trips==0.2)

# facet by species. model options are including space, effort, both, neither (color)
dat$Covariate = paste(dat$spatial, dat$effort)
dat$Covariate[which(dat$Covariate==" ")] = "Neither"
dat$Covariate[which(dat$Covariate=="SP EFF")] = "Space, Effort"
dat$Covariate[which(dat$Covariate=="SP ")] = "Space"
dat$Covariate[which(dat$Covariate==" EFF")] = "Effort"

# --------------------------------------------------------------
# v3: only Ratio, GAM-Tweedie, and RF (no delta models)
#   - add "AVERAGE" last panel summarizing across all species
# add "AVERAGE" as a 'species' to put in bottom right panel
sp.labs <- levels(dat$species)
dat$species <- as.character(dat$species)
dat.allspecies <- dat
dat.allspecies$species = "AVERAGE"
dat.all <- rbind(dat, dat.allspecies)
dat.all$species <- factor(dat.all$species, levels = c(sp.labs,"AVERAGE"))

filter_dat = group_by(dat.all, species, model, Covariate) %>% 
  mutate(rmse = ifelse(rmse > 10, NA, rmse)) %>% 
  ungroup() %>% 
  group_by(species) %>% 
  mutate(max_rmse = max(rmse,na.rm=T))
filter_dat$delta <- as.numeric(grepl("Delta", filter_dat$model))
filter_dat$delta <- factor(filter_dat$delta,labels=c("Non-delta","Delta"))

dat.spatial.eff <- filter(filter_dat, Covariate=="Space, Effort") %>% mutate(medRatio = NA) %>%
                    filter(model %in% c("Ratio", "GAM", "RF"))
dat.ratio <- filter(filter_dat, model=="Ratio") %>% group_by(species) %>% mutate(medRatio = median(rmse))
dat.plot <- rbind(dat.spatial.eff, dat.ratio)
dat.plot$model <- factor(dat.plot$model, levels = c("Ratio", "GAM", "RF"))
dat.plot$species <- factor(dat.plot$species, levels = c(sp.labs,"AVERAGE"))

print(dat.plot %>% #group_by(species, model) %>%
  ggplot(aes(model, rmse, fill = model)) + 
  # ggplot(aes(model, rmse, fill = model, alpha=delta)) + 
    # geom_boxplot(outlier.shape = NA) + 
    geom_violin(draw_quantiles = c(0.5)) +
    geom_hline(aes(yintercept = medRatio, group = species), linetype = 2) +
    facet_wrap(~ species) + 
    theme_sleek() + 
    xlab("Model") + 
    ylab("Normalized RMSE") + 
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
          # strip.text.x = element_text(face=c(rep("plain",15),"bold"))) + 
    # strip.text.x = element_text(face="bold")) + 
    coord_cartesian(ylim = c(0,1)) +
  scale_fill_manual(name="Model", 
                    labels=c("Ratio", "GAM Tweedie", "RF Total"),
                    values=c("grey","#E69F00","#56B4E9")))

# # overall stats
# library(tidyr)
# dat.plot$model <- as.character(dat.plot$model)
# tmp <- as.data.frame(dat.plot %>% filter(species=="OVERALL")) %>% select(model, rmse) %>% 
#             group_by(model) %>% mutate(id = row_number()) %>% spread(model, rmse) %>% select(-id) %>%
#             mutate(GAM.Ratio=GAM-Ratio, RF.Ratio=RF-Ratio) %>%
#             mutate(GAM.Ratio.perc=GAM.Ratio/Ratio, RF.Ratio.perc=RF.Ratio/Ratio)
# round(apply(tmp,2,median,na.rm=T),2)
# 
# #            GAM          Ratio             RF      GAM.Ratio       RF.Ratio 
# #           0.21           0.22           0.16          -0.01          -0.06 
# # GAM.Ratio.perc  RF.Ratio.perc 
# #          -0.05          -0.27 
```

