---
title: "ggplot2 examples"
author: "Hao Ye"
date: "`r format(Sys.time(), '%d %B, %Y')`"
css: styles.css
output:
  html_document:
    toc: false
    number_sections: true
---

```{r setup, include = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)
```

# Basic ggplot syntax

The `ggplot2` package operates by building a plot in layers, corresponding to different aspects of a plot. This modular approach makes it simple to flexibly modify aspects of a plot, but with a steeper learning curve for starting out.

First, let's load the package and some sample data.
```{r ggplot setup}
library(ggplot2)
data(diamonds)
```

Now, let's create a simple ggplot object and define several layers. I recommend using the standard `ggplot` function rather than `qplot` to take advantage of the modular properties of using ggplot.

```{r simple data and mapping}
p <- ggplot(data = diamonds, 
            mapping = aes(x = carat, 
                          y = price, 
                          color = cut))
summary(p)
```

Our object `p` now includes the data, and a mapping that describes the x-variable, the y-variable, and the variable that will specify the color.

# Some sample plot types

```{r scatterplot}
p + geom_point()
```

```{r lineplot}
p + geom_line()
```

```{r boxplot}
p + geom_boxplot()
```

```{r 2d histogram}
p + stat_bin2d()
```

```{r 2d hexplot}
p + geom_hex()
```

# Modifying attributes

We can of course modify various aspects of the plots.

```{r scatterplot ii}
p + geom_point() + 
    xlab("Size (carat)") + 
    ylab("Price (USD?)") +
    coord_cartesian(xlim = c(0, 3), ylim = c(0, 10000))
```

```{r scatterplot iii}
p + geom_point() + 
        coord_flip() + 
        scale_color_manual(values = rainbow(5))
```

```{r scatterplot iiii}
p + geom_point() + 
        facet_grid(clarity ~ .) + 
        theme(panel.grid.major = element_blank(),
              panel.grid.minor = element_blank(),
              axis.text = element_text(color = "black"), 
              panel.background = element_rect(color = "black", fill = NA))
```


# More complex plots

The modular nature also makes it straightforward to combine multiple types of plots or to add features. For example, there are several built-in functions to do common summarizations:

```{r scatter with loess regression}
p + geom_point() + geom_smooth(method = "loess")
```

Of course, it is also possible to plot your own analyses:
```{r scatter with custom predictions}
demo_model <- glm(price ~ carat * cut, data = diamonds)
sample_df <- diamonds[sample(1:NROW(diamonds), 1000), c("cut", "carat")]
sample_df$price <- predict(demo_model, sample_df)
p + geom_point() + geom_line(data = sample_df, size = 2)
```

