---
title: "Github Intro"
author: "Brian Stock"
date: "November 1, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## What is git? What is Github?

Git is a version control system, "a tool that keeps track of these changes for us and helps us version and merge our files." <http://swcarpentry.github.io/git-novice/01-basics/>

"GitHub is a code hosting platform for version control and collaboration. It lets you and others work together on projects from anywhere." <https://guides.github.com/activities/hello-world/>

"The **sweet spot for this kind of training is therefore the first years of
graduate school**. At that point, students have *time to learn* (at least,
more time than they’ll have once they’re faculty) and *real problems
of their own* that they want to solve." [Wilson 2016](https://f1000research.com/articles/3-62/v2)

### Example Github repositories

* Brian's repositories: <https://github.com/brianstock>
* SIO-BUGS repository: <https://github.com/SIO-BUG/BUG-Resources>.
* SIO-R-Users repository: <https://github.com/ha0ye/SIO-R-Users>.

### Unlimited private repos for students

While a student you can request a ["developer pack"](https://education.github.com/pack) that gives you unlimited private repositories for chapters/papers in progress.

## Let's get started

1. Create Github account
2. Clone/download SIO-R-Users repository

```{r eval=F}
git status
git pull origin master
git status
```

3. Create new folder and file
4. Commit changes

```{r eval=FALSE}
git add -A
git commit -m “message”
git status
```

5. Push to Github

```{r eval=F}
git push origin master
```

6. Check online, see changes

## Future workflow: adding a presentation to R-Users Github

```{r eval=F}
...cd to SIO-R-Users folder...
git pull origin master
...make local changes...
git add -A
git commit -m “message”
git push origin master
```

## Existing resources

In case this totally flops or we can't figure out your machine, [Software Carpentry](http://software-carpentry.org/lessons/) has a great set of tutorials for git/Github (as well as R and other things computing): 

* [Git intro (3 hour workshop)](http://swcarpentry.github.io/git-novice/reference/)
* [Recorded git intro workshop](https://www.youtube.com/watch?v=hKFNPxxkbO0)

Github also has nice tutorials for getting started: <https://guides.github.com/>.
