---
title: "How fast is R?"
author: "Hao Ye"
date: "`r format(Sys.time(), '%d %B, %Y')`"
css: styles.css
output:
  html_document:
    toc: false
    number_sections: true
---

# Everything I learned in Algorithms Class is Useless

- Datasets can be large, but rarely "humongous"
- R code is often run just a few times or even once
- Computer time matters, but human time (sometimes even grad students!) usually matters more
- R is usually "fast enough" (if you make an effort to use built-in functions instead of loops)

# Simple Ways to Time Your Code

## Setup

```{r slowish example, cache = TRUE}

# make some data (note data.frame type)
data <- as.data.frame(
    matrix(rnorm(400000 * 150, mean = 5), ncol = 15)
)

# define a function to normalize the columns to be mean 0
normCols <- function(d) {
    means <- apply(d, 2, mean) # compute means over each column
    
    for(i in seq_along(means)) {
        d[, i] <- d[, i] - means[i]
    }
    
    # to not have to deal with the output, we return nothing, though normally, 
    # we would return the object d
    return()
}
```

## system.time()

Just wrap your code or function calls in system.time.

```{r using system.time, cache = TRUE}
system.time(normCols(data))
```

## proc.time

Alternatively, you can get more manual control using \code{proc.time()}, which returns the amount of time the currently running R process has already taken.

```{r using proc.time, cache = TRUE}
start_time <- proc.time() # record when we started our code

normCols(data)

print(proc.time() - start_time) # output elapsed difference
```

## proc.time (advanced)

```{r using proc.time ii, cache = TRUE}
start_time <- proc.time()

normCols(data)

message("Elapsed time was ", (proc.time() - start_time)[3], " seconds.")
```

## profvis

Using the "profvis" package, we can have a nice visualization of the output from the built-in R profiler:

```{r using profvis, cache = TRUE}
library(profvis)

profvis(normCols(data))
```

If we include the function definition, the display will even track down the elapsed time to individual lines of code. Note that we wrap our code block in curly braces, {}.

```{r using profvis ii, cache = TRUE}
profvis({
    normCols <- function(d) {
        means <- apply(d, 2, mean) # compute means over each column
        
        for(i in seq_along(means)) {
            d[, i] <- d[, i] - means[i]
        }
        
        # to not have to deal with the output, we return nothing, though normally, 
        # we would return the object d
        return()
    }
    normCols(data)
})
```

Since most of the time was used in the apply function, let's try a couple different versions to do the same calculation.

```{r using profvis iii, cache = TRUE}
library(Rcpp)

profvis({
    # means <- apply(data, 2, mean)
    means <- colMeans(data)
    means <- lapply(data, mean)
    means <- vapply(data, mean, numeric(1))
    
    # means <- lapply(data, function(v) {
    #     total <- 0
    #     for(i in seq_along(v)) {total <- total + v[i]}
    #     return(total / length(v))
    # })
    
    means <- lapply(data, function(v) {
        return(sum(v) / length(v))
    })
    
    cppFunction( "double cpp_sum(const NumericVector v) {
                 double total = 0;
                 for(size_t i = 0; i < v.size(); ++i) {total += v[i];}
                 return(total);
                 }")
    
    means <- lapply(data, function(v) {
        total <- cpp_sum(v)
        return(total / length(v))
    })
})

library(plyr)

means <- ddply(data, mean)

```

# Links
- [More info on profvis](rstudio.github.io/profvis)

